Automated topic naming to support cross-project analysis of software maintenance activities

### 2011-01-21 msr/msr2011.tex
moved discussion of findings to discussion section ensmallened the author emails revised title updated sources of error

\author{
\alignauthor 
Abram Hindle\\
   \affaddr{Dept.of Computer Science}\\
   \affaddr{University of California, Davis} \\
   \affaddr{Davis, CA, USA}\\
   \email{abram@softwareprocess.es}
\alignauthor  Neil A. Ernst\\
   \affaddr{Dept. of Computer Science}\\
   \affaddr{University of Toronto}\\
   \affaddr{Toronto, Ontario, CANADA}\\
   \email{nernst@cs.toronto.edu}
\and
\alignauthor Michael W. Godfrey\\
   \affaddr{David Cheriton School of Computer Science}\\
   \affaddr{University of Waterloo}\\
   \affaddr{Waterloo, Ontario, CANADA}\\
   \email{migod@uwaterloo.ca}
\alignauthor John Mylopoulos \\
   \affaddr{Dept. Information Eng. and Computer Science}\\
   \affaddr{University of Trento}\\
   \affaddr{Trento, ITALY}\\
   \email{\normalsize jm@disi.unitn.it}
}

%\title{What's in a name? Automated topic naming of software maintenance activities}
\title{Automated topic naming to support analysis of software maintenance activities}
\maketitle
%\thispagestyle{empty}

@@

Topic modeling is a machine learning technique which creates multinomial distributions over words found in a text corpus. Topic models are useful in software engineering because they can identify, from text corpora such as source code, change logs, email, and others, co-occuring words. For example, a topic extracted from a set of changelogs might be ``listener change remove add fire''. A topic model gives developers an overview of where activity is occurring and gives managers or maintainers a sense of project history. A major challenge is to accurately identify the label for a topic (in the previous case, perhaps ``event-handling''). Current approaches are manual: they rely on human expertise to recognize the appropriate label. Furthermore, to date the results of topic modeling are project specific, and not generalizable.

@@

\subsection{Annotation observations}
We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, likely because concurrent code is prone to bugs such as race conditions. %are 
% often  was troublesome. 
\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration management. 
these kinds of changes are slightly related to \emph{maintainability}. 
A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files. In these cases we assigned the \emph{None} class.

@@

%XXX can just remove this
%\subsection{Effectiveness}

@@

% can always be cut down
Our work faced multiple threats to validity and we attempted to address many of them.

\emph{Construct validity} -- we used only commit messages rather than mail or bug tracker messages. 
Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
We used mailing-lists to verify the purpose behind some events.
We also chose our taxonomy and the data to study. We did not show these results to the stakeholders.
We only studied database systems.

\emph{Internal validity} -- includes inter-rater reliability and our lack of rival explanation building. 
We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
We did not attempt to match our results to any particular model or theory.

\emph{External validity} was addressed by our use of multiple case studies, but there are still issues. 
Our data originated from OSS database projects and thus might not be applicable to commercially developed software. 
Furthermore, our analysis techniques rely on a project's use of meaningful commit messages. The OSS projects investigated were database systems, thus 
our results might not generalize to other domains.
% \todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}

% in Case Study Research: Design Case Studies
\emph{Reliability} was addressed by each annotator following the same protocol and using the same annotations. 
Unfortunately only two annotators were used; their annotations could be biased as we did not analyze for inter-rater reliability.

>>>>

\author{
\alignauthor 
Abram Hindle\\
   \affaddr{Dept. of Computer Science}\\
   \affaddr{University of California, Davis} \\
   \affaddr{Davis, CA, USA}\\
   \email{\large abram@softwareprocess.es}
\alignauthor  Neil A. Ernst\\
   \affaddr{Dept. of Computer Science}\\
   \affaddr{University of Toronto}\\
   \affaddr{Toronto, Ontario, CANADA}\\
   \email{\large nernst@cs.toronto.edu}
\and
\alignauthor Michael W. Godfrey\\
   \affaddr{David Cheriton School of Computer Science}\\
   \affaddr{University of Waterloo}\\
   \affaddr{Waterloo, Ontario, CANADA}\\
   \email{\large migod@uwaterloo.ca}
\alignauthor John Mylopoulos \\
   \affaddr{Dept. Information Eng. and Computer Science}\\
   \affaddr{University of Trento}\\
   \affaddr{Trento, ITALY}\\
   \email{\large jm@disi.unitn.it}
}

%\title{What's in a name? Automated topic naming of software maintenance activities}
\title{Automated topic naming to support cross-project analysis of software maintenance activities}
\maketitle
%\thispagestyle{empty}

@@

Topic modeling is a machine learning technique which creates multinomial distributions over words from a text corpus. Topic models are useful in software engineering because they can identify co-occuring words from text corpora such as source code, commit comments, email, and others. For example, a topic extracted from a set of commit logs might be\emph{ ``listener change remove add fire''}. A topic model is a summary of the key concepts in the corpus. Among other uses, it can quickly give developers an overview of where activity is occurring and gives managers or maintainers a sense of project history. A major challenge is to accurately identify the summary label for a topic (in the previous case, perhaps \emph{``event-handling''}). 
Current approaches are manual: they rely on human expertise to recognize the appropriate label. Furthermore, to date the results of topic modeling are project specific, and not generalizable. This paper presents an approach, labelled topic extraction, which addresses these two problems.

@@

@@

%XXX can just remove this
%\subsection{Effectiveness}
\subsection{Annotation observations}
We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, likely because concurrent code is prone to bugs such as race conditions. %are 
% often  was troublesome. 
\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration management. 
these kinds of changes are slightly related to \emph{maintainability}. 
A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files. In these cases we assigned the \emph{None} class.

@@

% can always be cut down
%Our work faced multiple threats to validity and we attempted to address many of them.

\emph{Construct validity} -- we used only commit messages rather than mail or bug tracker messages. Since we required a seamless record, it was not possible to obtain other data sources. Possibly they would have influenced our results, but there would be a degree of correlation between the corpora.
%Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
%We used mailing-lists to verify the purpose behind some events.
Our taxonomy for software NFRs is subject to dispute, but seems generally accepted. Finally, there are exogenous sources, such as in-person discussions, which we did not access, an omission as noted in \cite{aranda09icse}. %We did not show these results to the stakeholders.

\emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 
We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
We did not attempt to match our results to any particular model or theory, as such theories are sparse in

\emph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
Our data originated from OSS database projects and thus might not be applicable to commercially developed software. 
Furthermore, our analysis techniques rely on a project's use of meaningful commit messages. The OSS projects investigated were database systems, thus 
our results might not generalize to other domains. %We only studied database systems.
% \todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}

% in Case Study Research: Design Case Studies
\emph{Reliability} -- each annotator followed the same protocol and used the same annotations. 
However, only two annotators were used; their annotations could be biased as we did not analyze for inter-rater reliability.
